---
title: "Getting Started with chartjs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with chartjs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `chartjs` package provides R bindings for the Chart.js JavaScript library, allowing you to create interactive charts directly from R. This vignette will walk you through the basics of creating different types of charts.

```{r setup}
library(chartjs)
```

## Basic Bar Chart

Let's start with a simple bar chart:

```{r bar_chart}
# Create sample data
data <- data.frame(
  categories = c("Red", "Blue", "Yellow", "Green", "Purple", "Orange"),
  values = c(12, 19, 3, 5, 2, 3)
)

# Create bar chart
chartjs_bar(data, x = "categories", y = "values")
```

## Line Chart

For time series or continuous data, line charts are ideal:

```{r line_chart}
# Create sample data
months <- c("January", "February", "March", "April", "May", "June", "July")
sales <- c(65, 59, 80, 81, 56, 55, 40)
expenses <- c(28, 48, 40, 19, 86, 27, 90)

data <- data.frame(
  month = months,
  sales = sales,
  expenses = expenses
)

# Create line chart with multiple series
chartjs_line(data, x = "month", y = c("sales", "expenses"))
```

## Pie Chart

For showing proportions, use pie charts:

```{r pie_chart}
# Create sample data
data <- data.frame(
  browser = c("Chrome", "Firefox", "Safari", "Edge", "Other"),
  usage = c(55, 25, 10, 7, 3)
)

# Create pie chart
chartjs_pie(data, labels = "browser", values = "usage")
```

## Scatter Plot

For exploring relationships between two variables:

```{r scatter_plot}
# Create sample data
set.seed(123)
n <- 50
data <- data.frame(
  height = rnorm(n, 170, 10),
  weight = rnorm(n, 70, 15)
)

# Create scatter plot
chartjs_scatter(data, x = "height", y = "weight")
```

## Customizing Charts

You can customize charts by passing options:

```{r custom_chart}
# Custom options
custom_options <- list(
  plugins = list(
    title = list(
      display = TRUE,
      text = "Monthly Sales Data"
    ),
    legend = list(
      position = "bottom"
    )
  ),
  scales = list(
    y = list(
      beginAtZero = TRUE,
      title = list(
        display = TRUE,
        text = "Sales ($1000s)"
      )
    ),
    x = list(
      title = list(
        display = TRUE,
        text = "Month"
      )
    )
  )
)

# Create customized bar chart
chartjs_bar(data, x = "month", y = "sales", options = custom_options)
```

## Integration with Shiny

The `chartjs` package integrates seamlessly with Shiny applications:

```{r shiny_example, eval=FALSE}
library(shiny)
library(chartjs)

ui <- fluidPage(
  titlePanel("Chart.js in Shiny"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("chart_type", "Chart Type:",
                  choices = c("Bar" = "bar", "Line" = "line", "Pie" = "pie")),
      
      actionButton("update", "Update Chart")
    ),
    
    mainPanel(
      chartjsOutput("myChart")
    )
  )
)

server <- function(input, output, session) {
  # Sample data
  data <- reactive({
    data.frame(
      categories = c("A", "B", "C", "D"),
      values = sample(10:50, 4)
    )
  })
  
  # Render chart
  output$myChart <- renderChartjs({
    chartjs(data(), type = input$chart_type, x = "categories", y = "values")
  })
  
  # Update chart dynamically
  observeEvent(input$update, {
    new_data <- data.frame(
      categories = c("A", "B", "C", "D"),
      values = sample(10:50, 4)
    )
    
    chartjs_proxy("myChart") %>%
      chartjs_update_data(new_data, x = "categories", y = "values")
  })
}

shinyApp(ui = ui, server = server)
```

## Next Steps

This vignette covered the basics of using the `chartjs` package. For more advanced features, explore:

- Animation options
- Plugin system integration  
- Custom themes and styling
- Advanced Shiny interactions
- Real-time data updates

Check out the package documentation for more examples and detailed API reference.